Truthfulness Verification System

### 2011-04-27 proposal/paper.tex
reorganized the sections and corrected sectio subsection

\section{Progress Summary}

{\em Mar 29 to Apr 12}
See page~[\pageref{Mar29toApr12}] of section~\ref{sec:UseOfWikipedia}
\begin{itemize}
\item Extraction of meaningful entities(noun phrase) in topic unit and dount unit by checking all combination of content words
\item Ranking based on co-occurence of content words in the interesction of bag of words generated from wiki-articles. 
\item Ranking based on co-occurence of content words in the sentences generated from wiki-articles. 
\end{itemize}

{\em Mar 9 to Mar 29}
\begin{itemize}
\item Formulated the initial steps for feature extraction for each doubt unit using Wikipedia. See page~[\pageref{March9to29}] of section~\ref{sec:UseOfWikipedia}
\end{itemize} 
{\em Feb 23 to Mar 9} 
\begin{itemize}
\item Formulated the procedure of truthfulness verification in a more structured way
\item As a result of the procedure formulation, the sections Step-by-step Algorithm for Truthfulness and Verification Examples of Truthfulness Verification are written in a complete revised way compared to the proposal
\item Apart from getting out feet wet with the MQL api, the daunting task is to develop generalized framework that is flexible enough to fit the wide variety of sentences into the rigid structure of Freebase. The learning cure is steep, given the richness and massive size of documentation. Though we are groping in dark currently, we are hopeful that the pieces of the puzzle with soon fit in.
\item Looked into applicability of POS tagger in the context of our problem and found it useful for constructing freebase queries
\item Got some initial promising results by exploring different options such as Wikipedia, answer.com beyond Freebase ontologies. 
\item Worked on porting of existing T-verifier code to linux. The original T-verifier code is for windows only and not generic enough at this moment. We also contacted with the student author about the problems we are facing to port the code to linux. We hope to resolve it soon.
\end{itemize}

@@

\section {Proposed System}

\begin{figure}
  \centering
  \subfloat[System
  overview]{\label{fig:System}\includegraphics[width=0.85\textwidth]{system.pdf}}
  \caption{System overview}
  \label{fig:system}
\end{figure}

\subsection{Description of extraction algorithm}

@@

\subsubsection{Freebase} 
Freebase has information about approximately 20 million Topics, each one having a
unique Id, which can help distinguish multiple entities which have similar names,
such as Henry Ford the industrialist vs Henry Ford the footballer. Most of the
topics are associated with one or more types\cite{freebasetype} (such as
people, places, books, films, etc) and may have additional properties like "date of birth" for a person
or latitude and longitude for a location. Freebase not only contains data from
the Wikipedia but also other sources; users can submit data to the Freebase
datastore and expand it in richness. We tinkered with the api\cite{freebaseapi}
and it appeared to be the most viable starting point for the project.

@@

\subsubsection{Dbpedia}
DBpedia is a similar project to Freebase, but it focuses mainly on the content
available from Wikipedia. It scores in being precisely importing the data from
the info boxes in Wikipedia pages, but at this stage it does not seem to be
offering anything additional over Freebase \cite{freebasedbpedia}. We are yet to
explore its programmatic interface \cite{dbpediaapi}.

\subsubsection{Yago}
YAGO is a semantic knowledge base with over 900,000 entities (like persons,
organizations, cities, etc.) and uses Wikipedia and Wordnet as its main source of
information. We are yet to explore the programmatic interfaces it provides and
how we can use it for the project.

@@

\section{Use of Wikipedia}
\label{sec:UseOfWikipedia}
We have tried two approaches for using the Wikipedia. In both we utilize the content words to identify articles and there upon treat them separately. In the first approach we treat the articles as bag of words and do not consider the sentence structure and in the next approach we retain the logical structure of each sentence and try to infer from word overlaps.
%When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

\subsection{Examples}
Examples of the algorithm described above is discussed below. The algorithm 2 produce perfect result for three of the five sentences. But it struggles with the remaining two sentences. However, the algorithm 2 primarily augment the algorithm 1 and add strength to the results from algorithm 1. Hence they work better together and produce overall better output.

\subsubsection{Tom Hanks was the lead actress  int move 'Sleepless in Seattle'}
The alternative units for this sentence are "Tom Hanks", "Meg Ryan", "Hanks Meg", "Nora", and "Sam". Here the sentence containing "Tom Hanks" contain the word actor but not actress. So the algorithm classify it as false. But the sentence containing "Meg Ryan" has the word actress and hence classify it as true. For other alternative units, we don't even get any sentences containing the alternative unit. So the algorithm produce an unambiguous single true statement for this sentence.

\subsubsection{Les Paul invented the electric guitar}
The alternative units for this sentence are "Les Paul", "rickenbacker", "Gibson", "Leo", and "Lloyd". Now there is inherent ambiguity about the true inventor of the electric guitar. On reading the articles related above, there is no information that Rickenbacker actually inventing the electric guitar. He was a founding member of the Rickenbacker company that produces electric guitar. Les Paul on the other hand was a pioneer in designing and developing of what a prototype. Moreover, some of the categories attached with the Les Paul page are American musical instrument makers, Guitar makers, Inventors of musical instruments, National Inventors Hall of Fame inductees. In comparison Adolph Rickenbaker page has only one guitar related category - Guitar stubs, meaning the article is stuck in content.

\subsubsection{Luna 2 is the first spacecraft on the moon}
The alternative units for this sentence are "Luna 2", "Apollo", "Land", "Landing", and "Nasa". Our algorithm returns true for both Luna 2 and Apollo and return false for all others. Now, both Luna 2 and Apollo are in reality true considering the facts that Luna 2 was the first unmanned spacecraft and Apollo was the first manned spacecraft. Hence, we can't say one of the Luna 2 or Apollo as false if farther information is not given. So algorithm 2 can produce correct result for this sentence also.

\subsubsection{English is the Primary Language of the Philippines}
The alternative units for this sentence are "English", "Filipino", "Spanish", "Education", and "History". Our current algorithm struggle with this sentence because of occurrence of content words and alternative units in the same sentence for all the alternative units. So it classify all of them as true and assign same weight to all of them. However, as our algorithm 1 can rank the alternative units in correct order, merged rank from algorithm 1 and algorithm 2 is still correct and produce overall correct result.

\subsubsection{Sullivan is the fastest swimmer in the world}
The alternative units for this sentence are "Sullivan", "Michael phelps", "ALexander", "Long", and "Ernest". Similar to the previous sentence, the algorithm 2 can not produce good ranking for for this sentence. However, our algorithm 1 produces correct ranking and together they produce overall correct result.

\section{Rank merging}
This ranking procedure bolsters the result in the previous algorithm always without any exception. On observing the five examples above, we expect a combination of the two algorithms will give us a proper ranking of the alternate units for proper truth verification.

>>>>

%\section{Progress Summary}
%
%
%{\em Mar 29 to Apr 12}
%See page~[\pageref{Mar29toApr12}] of section~\ref{sec:UseOfWikipedia}
%\begin{itemize}
%\item Extraction of meaningful entities(noun phrase) in topic unit and dount unit by checking all combination of content words
%\item Ranking based on co-occurence of content words in the interesction of bag of words generated from wiki-articles. 
%\item Ranking based on co-occurence of content words in the sentences generated from wiki-articles. 
%\end{itemize}
%
%
%{\em Mar 9 to Mar 29}
%\begin{itemize}
%\item Formulated the initial steps for feature extraction for each doubt unit using Wikipedia. See page~[\pageref{March9to29}] of section~\ref{sec:UseOfWikipedia}
%\end{itemize} 
%{\em Feb 23 to Mar 9} 
%\begin{itemize}
%\item Formulated the procedure of truthfulness verification in a more structured way
%\item As a result of the procedure formulation, the sections Step-by-step Algorithm for Truthfulness and Verification Examples of Truthfulness Verification are written in a complete revised way compared to the proposal
%\item Apart from getting out feet wet with the MQL api, the daunting task is to develop generalized framework that is flexible enough to fit the wide variety of sentences into the rigid structure of Freebase. The learning cure is steep, given the richness and massive size of documentation. Though we are groping in dark currently, we are hopeful that the pieces of the puzzle with soon fit in.
%\item Looked into applicability of POS tagger in the context of our problem and found it useful for constructing freebase queries
%\item Got some initial promising results by exploring different options such as Wikipedia, answer.com beyond Freebase ontologies. 
%\item Worked on porting of existing T-verifier code to linux. The original T-verifier code is for windows only and not generic enough at this moment. We also contacted with the student author about the problems we are facing to port the code to linux. We hope to resolve it soon.
%\end{itemize}

@@

\section{Proposed System}
\label{sec:UseOfWikipedia}
We have tried two approaches for using the Wikipedia. In both we utilize the content words to identify articles and there upon treat them separately. In the first approach we treat the articles as bag of words and do not consider the sentence structure and in the next approach we retain the logical structure of each sentence and try to infer from word overlaps.
%When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

\paragraph{Examples}
Examples of the algorithm described above is discussed below. The algorithm 2 produce perfect result for three of the five sentences. But it struggles with the remaining two sentences. However, the algorithm 2 primarily augment the algorithm 1 and add strength to the results from algorithm 1. Hence they work better together and produce overall better output.

\paragraph{Tom Hanks was the lead actress  int move 'Sleepless in Seattle'}
The alternative units for this sentence are "Tom Hanks", "Meg Ryan", "Hanks Meg", "Nora", and "Sam". Here the sentence containing "Tom Hanks" contain the word actor but not actress. So the algorithm classify it as false. But the sentence containing "Meg Ryan" has the word actress and hence classify it as true. For other alternative units, we don't even get any sentences containing the alternative unit. So the algorithm produce an unambiguous single true statement for this sentence.

\paragraph{Les Paul invented the electric guitar}
The alternative units for this sentence are "Les Paul", "rickenbacker", "Gibson", "Leo", and "Lloyd". Now there is inherent ambiguity about the true inventor of the electric guitar. On reading the articles related above, there is no information that Rickenbacker actually inventing the electric guitar. He was a founding member of the Rickenbacker company that produces electric guitar. Les Paul on the other hand was a pioneer in designing and developing of what a prototype. Moreover, some of the categories attached with the Les Paul page are American musical instrument makers, Guitar makers, Inventors of musical instruments, National Inventors Hall of Fame inductees. In comparison Adolph Rickenbaker page has only one guitar related category - Guitar stubs, meaning the article is stuck in content.

\paragraph{Luna 2 is the first spacecraft on the moon}
The alternative units for this sentence are "Luna 2", "Apollo", "Land", "Landing", and "Nasa". Our algorithm returns true for both Luna 2 and Apollo and return false for all others. Now, both Luna 2 and Apollo are in reality true considering the facts that Luna 2 was the first unmanned spacecraft and Apollo was the first manned spacecraft. Hence, we can't say one of the Luna 2 or Apollo as false if farther information is not given. So algorithm 2 can produce correct result for this sentence also.

\paragraph{English is the Primary Language of the Philippines}
The alternative units for this sentence are "English", "Filipino", "Spanish", "Education", and "History". Our current algorithm struggle with this sentence because of occurrence of content words and alternative units in the same sentence for all the alternative units. So it classify all of them as true and assign same weight to all of them. However, as our algorithm 1 can rank the alternative units in correct order, merged rank from algorithm 1 and algorithm 2 is still correct and produce overall correct result.

\paragraph{Sullivan is the fastest swimmer in the world}
The alternative units for this sentence are "Sullivan", "Michael phelps", "ALexander", "Long", and "Ernest". Similar to the previous sentence, the algorithm 2 can not produce good ranking for for this sentence. However, our algorithm 1 produces correct ranking and together they produce overall correct result.

\subsection{Rank merging}
This ranking procedure bolsters the result in the previous algorithm always without any exception. On observing the five examples above, we expect a combination of the two algorithms will give us a proper ranking of the alternate units for proper truth verification.

\section {Alternative methods that are explored}

%\begin{figure}
%  \centering
%  \subfloat[System
%  overview]{\label{fig:System}\includegraphics[width=0.85\textwidth]{system.pdf}}
%  \caption{System overview}
%  \label{fig:system}
%\end{figure}
%
%
%\subsection{Description of extraction algorithm}

@@

\subsection{Freebase} 
Freebase has information about approximately 20 million Topics, each one having a
unique Id, which can help distinguish multiple entities which have similar names,
such as Henry Ford the industrialist vs Henry Ford the footballer. Most of the
topics are associated with one or more types\cite{freebasetype} (such as
people, places, books, films, etc) and may have additional properties like "date of birth" for a person
or latitude and longitude for a location. Freebase not only contains data from
the Wikipedia but also other sources; users can submit data to the Freebase
datastore and expand it in richness. We tinkered with the api\cite{freebaseapi}
and it appeared to be the most viable starting point for the project.

@@

\subsection{Dbpedia}
DBpedia is a similar project to Freebase, but it focuses mainly on the content
available from Wikipedia. It scores in being precisely importing the data from
the info boxes in Wikipedia pages, but at this stage it does not seem to be
offering anything additional over Freebase \cite{freebasedbpedia}. We are yet to
explore its programmatic interface \cite{dbpediaapi}.

\subsection{Yago}
YAGO is a semantic knowledge base with over 900,000 entities (like persons,
organizations, cities, etc.) and uses Wikipedia and Wordnet as its main source of
information. We are yet to explore the programmatic interfaces it provides and
how we can use it for the project.

@@
